---
title: "DataSHIELD Use-case: Distributed non-disclosive validation of predictive models by a modified ROC-GLM"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

- [About the repository](#about-the-repository)
    - [Structure of the repository](#structure-of-the-repository)
- [Setup](#setup)
    - [Install packages](#install-packages)
    - [Generate data and fit model](#generate-data-and-fit-model)
    - [Install package on DataSHIELD and upload data](#install-package-on-datashield-and-upload-data)
- [Analysis](#inspect-results-using-the-docker)
    - [Log into DataSHIELD test server](#log-into-datashield-test-server)
    - [Push and predict](#push-and-predict)
    - [Analyse calibration of the predictions](#analyse-calibration-of-the-predictions)
    - [Evaluate the model using ROC analysis](#evaluate-the-model-using-roc-analysis)
    - [Cross check on pooled test data](#cross-check-on-pooled-test-data)
- [Log out from DataSHIELD servers](#log-out-from-datashield-servers)


```{r, include=FALSE}
source(here::here("R/generate-data.R"))
source(here::here("R/create-model.R"))
source(here::here("R/upload-data.R"))
source(here::here("R/install-ds-packages.R"))

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/"
)

library(ggplot2)
library(ggsci)

### Theme setup:
## ==================================================================

# Set font (if available):
font = "Tinos"
sysfonts::font_add_google(font, font)
extrafont::loadfonts()
ft = extrafont::fonttable()

if (all(! grepl(font, x = ft$FamilyName))) {
  my_theme = theme_minimal()
} else {
  my_theme = theme_minimal(base_family = font)
}
# to determine width, use the latex package `layouts` and put
# `\printinunitsof{mm}\prntlen{\textwidth}` into the tex file.
textwidth = 148.92

#' Open pdf files using evince on linux machines
#' @param file to the pdf
evince = function(file) system(paste0("evince ", file, " &"))
```

<!-- README.md is generated from README.Rmd. Please edit that file -->

## About the repository

This repository contains a short use-case base on the three packages `dsPredictBase`, `dsCalibration`, and `dsROCGLM`. The main intend is to have a use-case to demonstrate how to distributively evaluate a model using the distributed [ROC-GLM](https://pubmed.ncbi.nlm.nih.gov/10877289/).

The following contains the preparation of test data and a test model as [setup](#setup) while the second part is the [analysis](#analysis).

Last time rendered: `r as.character(format(Sys.time(), "%H:%M - %d. %h %Y"))` by user `r Sys.getenv()[["USER"]]`
Autobuild: [![Render README](https://github.com/difuture-lmu/datashield-demo-survival/actions/workflows/render-readme.yaml/badge.svg)](https://github.com/difuture-lmu/datashield-demo-survival/actions/workflows/render-readme.yaml)

### Structure of the repository

- `R`:
    - `create-model.R`: Creates a [`ranger`](https://cran.r-project.org/web/packages/ranger/ranger.pdf) used for the use-case based on the data in `generate-data.R`
    - `generate-data.R`: Takes the data set `GBSG2` (see `?GBSG2` for a description) from the [`TH.data`](https://cran.r-project.org/web/packages/TH.data/index.html), splits it into trian and test using 60 - 40 % of the data, and furhter splits the 40 % for testing into 5 parts for the distributed setup.
    - `helper.R`: Helper functions to locally calculate the [ROC-GLM](https://pubmed.ncbi.nlm.nih.gov/10877289/) and compute confidence intervals etc.
    - `install-ds-packages.R`: Install the necessary packages (`ranger`, `dsPredictBase`, `dsCalibration`, and `dsROCGLM`) __at the DataSHIELD servers__.
    - `install-packages.R`: Install ncessary packages locally.
    - `upload-data.R` Creates a project at the DataSHIELD server and uploads the data created by `generate-data.R`.
- `data`: All data is stored here:
    - Train and test split of the GBSG2 data set (`data-train.csv` and `data-test.csv`).
    - The 5 splits of the `data-test.csv` for the servers (`SRV1.csv`, `SRV2.csv`, `SRV3.csv`, `SRV4.csv`, and `SRV5.csv`).
    - The model created by `create-model.R` (`mod.Rda`).
    - A csv file for logging each rendering (`log.csv`). This file can be used to get an overview about the important values and when each rendering was conducted. The main purpose is to show that the results are reproduced at each rendering.
    - The ROC-GLM of the last rendering (`roc-glm.Rda`).
- `figures`: Figures created by the rendering are placed here. These are the `.pdf` fuiles used in the publication but also the `.png` files of the README.
- `tables`: Tables created by the rendering are placed here.

## Setup

### Install packages

Install all packages locally:

```{r}
remotes::install_github("difuture-lmu/dsPredictBase", upgrade = "never")
remotes::install_github("difuture-lmu/dsCalibration", upgrade = "never")
remotes::install_github("difuture-lmu/dsROCGLM", upgrade = "never")
```

### Generate data and fit model

The used data is the `GBSG2` from the `TH.data` packages. For further details see the help page `?TH.data::GBSG2`. The task is to predict whether hormonal therapy shows an improvement w.r.t. survival time. The model we are using is a random forest from the `ranger` package. The following code uses the `GBSG2` data, splits it into train and test data with 60 % for training and 40 % for testing. The test data is further split into 5 parts that are uploaded to DataSHIELD and used to simulate the distributed setup.

```{r, eval=FALSE}
source(here::here("R/generate-data.R"))
source(here::here("R/create-model.R"))
```

### Install package on DataSHIELD and upload data

```{r, eval=FALSE}
source(here::here("R/upload-data.R"))
source(here::here("R/install-ds-packages.R"))
```

## Analysis

### Log into DataSHIELD test server

```{r}
library(DSI)
library(DSOpal)
library(dsBaseClient)

library(dsPredictBase)
library(dsCalibration)
library(dsROCGLM)

library(ggplot2)

builder = newDSLoginBuilder()

surl     = "https://opal-demo.obiba.org/"
username = "administrator"
password = "password"

datasets = paste0("SRV", seq_len(5L))
for (i in seq_along(datasets)) {
  builder$append(
    server   = paste0("ds", i),
    url      = surl,
    user     = username,
    password = password,
    table    = paste0("DIFUTURE-TEST.", datasets[i])
  )
}

## Get data of the servers:
conn = datashield.login(logins = builder$build(), assign = TRUE)
datashield.symbols(conn)
(ddim = ds.dim("D"))
```

### Push and predict

```{r}
## Load the pre-calculated logistic regression:
load(here::here("data/mod.Rda"))

## Push the model to the servers (upload takes ~11 Minutes):
t0 = proc.time()
pushObject(conn, obj = mod)
(t0 = proc.time() - t0)
datashield.symbols(conn)

## Time point:
which(ranger::timepoints(mod) >= 730)[1]

## Predict the model on the data sets located at the servers:
pfun = "ranger:::predict.ranger(mod, data = D)$survival[, 127]"
predictModel(conn, mod, "probs", predict_fun = pfun, package = "ranger")
datashield.symbols(conn)

# Because labels are flipped for the 0-1-setting we also calculate
# 1 - probs:
datashield.assign(conn, "pinv", quote(1 - probs))
```

### Analyse calibration of the predictions

```{r}
brier = dsBrierScore(conn, "D$valid", "pinv")
brier

cc = dsCalibrationCurve(conn, "D$valid", "pinv")
cc

gg_cal = plotCalibrationCurve(cc, size = 1)
gg_cal
```

```{r, include=FALSE}
# Save figure for paper
gg_cal_pub = gg_cal +
  scale_color_uchicago() +
  scale_fill_uchicago() +
  my_theme +
  theme(
    plot.title = element_text(size = 10),
    plot.subtitle = element_text(size = 9),
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 9),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    panel.border = element_rect(colour = "black", fill = NA, size = 0.5)
  )

ggsave(plot = gg_cal_pub,
  filename = here::here("figures/analysis-distr-calibration.pdf"),
  width = 0.7 * textwidth,
  height = textwidth * 0.4,
  units = "mm")

# evince(here::here("figures/analysis-distr-calibration.pdf"))
```

### Evaluate the model using ROC analysis

```{r}
# Get the l2 sensitivity
(l2s = dsL2Sens(conn, "D", "pinv"))
epsilon = 0.3
delta = 0.2

# Amount of noise added:
sqrt(2 * log(1.25 / delta)) * l2s / epsilon

# Calculate ROC-GLM
roc_glm = dsROCGLM(conn, "D$valid", "pinv", epsilon = epsilon,
  delta = delta, dat_name = "D", seed_object = "D$age")
roc_glm
roc_glm$ci

gg_distr_roc = plot(roc_glm)
gg_distr_roc
```

__Simple check if old (model from the last rendering) and new ROC-GLM are equal:__
```{r}
# Check if roc_glm object is the same for the new and the last run:
if (! file.exists(here::here("data/roc-glm.Rda"))) {
  saveRDS(roc_glm, file = here::here("data/roc-glm.Rda"))
} else {
  roc_glm_last_run = readRDS(here::here("data/roc-glm.Rda"))
  if (! identical(roc_glm_last_run, roc_glm))
    warning("Old and new ROC-GLM are not equal!")

  saveRDS(roc_glm, file = here::here("data/roc-glm.Rda"))
}
```

```{r, include=FALSE}
# Save figure for paper
df_auc = data.frame(lower = roc_glm$ci[1], upper = roc_glm$ci[2], auc = roc_glm$auc)

gg_roc_pub = plot(roc_glm, plot_ci = FALSE) +
  geom_errorbarh(data = df_auc,
    aes_string(y = "0.1", xmin = "lower", xmax = "upper"), height = 0.05) +
  geom_point(data = df_auc, aes_string(y = "0.1", x = "auc"), size = 3) +
  geom_point(data = df_auc, aes_string(y = "0.1", x = "auc"), size = 1, color = "white") +
  # annotate("text", x = df_auc$auc, y = 0.1, label = "AUC", vjust = -1, size = 3) +
  # annotate("text", x = df_auc$auc, y = 0.1, label = round(df_auc$auc, 2), vjust = 2, size = 2.5) +
  # annotate("text", x = df_auc$lower, y = 0.1, label = round(df_auc$lower, 2), hjust = 1.2, size = 3) +
  # annotate("text", x = df_auc$upper, y = 0.1, label = round(df_auc$upper, 2), hjust = -0.2, size = 3) +
  scale_color_uchicago() +
  scale_fill_uchicago() +
  my_theme +
  theme(
    plot.title = element_text(size = 10),
    plot.subtitle = element_text(size = 9),
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 9),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    panel.border = element_rect(colour = "black", fill = NA, size = 0.5)
  )

if (all(! grepl(font, x = ft$FamilyName))) {
  gg_roc_pub = gg_roc_pub +
    annotate("text", x = df_auc$auc, y = 0.1, label = "AUC", vjust = -1, size = 3) +
    annotate("text", x = df_auc$auc, y = 0.1, label = round(df_auc$auc, 2), vjust = 2, size = 2.5) +
    annotate("text", x = df_auc$lower, y = 0.1, label = round(df_auc$lower, 2), hjust = 1.2, size = 3) +
    annotate("text", x = df_auc$upper, y = 0.1, label = round(df_auc$upper, 2), hjust = -0.2, size = 3)
} else {
  gg_roc_pub = gg_roc_pub +
    annotate("text", x = df_auc$auc, y = 0.1, label = "AUC", vjust = -1, size = 3, family = font) +
    annotate("text", x = df_auc$auc, y = 0.1, label = round(df_auc$auc, 2), vjust = 2, size = 2.5, family = font) +
    annotate("text", x = df_auc$lower, y = 0.1, label = round(df_auc$lower, 2), hjust = 1.2, size = 3, family = font) +
    annotate("text", x = df_auc$upper, y = 0.1, label = round(df_auc$upper, 2), hjust = -0.2, size = 3, family = font)
}


ggsave(plot = gg_roc_pub,
  filename = here::here("figures/analysis-distr-roc-glm.pdf"),
  width = 0.5 * textwidth,
  height = textwidth * 0.5,
  units = "mm")

# evince(here::here("figures/analysis-distr-roc-glm.pdf"))
```


## Cross check on pooled test data

```{r}
#' Calculate TPR and FPRs to plot the empirical ROC curve
#'
#' @param labels (`integer()`) True labels as 0-1-coded vector.
#' @param scores (`numeric()`) Score values.
#' @return (`data.frame()`) of the TPR and FPRs.
simpleROC = function(labels, scores) {
  labels = labels[order(scores, decreasing = TRUE)]
  data.frame(
    TPR = cumsum(labels) / sum(labels),
    FPR = cumsum(! labels) / sum(! labels), labels)
}

# Load pooled test data and predict:
dat_test = read.csv(here::here("data/data-test.csv"), stringsAsFactors = TRUE)
probs = ranger:::predict.ranger(mod, data = dat_test)$survival[, 127]

# Calculate empirical AUC and compare with distributed ROC-GLM
auc = pROC::auc(dat_test$valid, 1 - probs)
c(auc_emp = auc, auc_distr_roc_glm = roc_glm$auc)

source(here::here("R/helper.R"))
(ci_emp = logitToAUC(pepeCI(toLogit(auc), 0.05, deLongVar(1 - probs, dat_test$valid))))


# Calculate TPR and FPR values and add to distributed ROC-GLM plot
plt_emp_roc_data = simpleROC(dat_test$valid, 1 - probs)

gg_roc_pooled = plot(roc_glm) +
  geom_line(data = plt_emp_roc_data, aes(x = FPR, y = TPR), color = "red")
gg_roc_pooled

# Calculate pooled brier score and calibration curve
brier_pooled = mean((dat_test$valid - (1 - probs))^2)
c(brier_pooled = brier_pooled, brier_distr = brier)

cc_pooled = calibrationCurve("dat_test$valid", "1 - probs", nbins = 10)

# Table of elements per server for the calibration curve:
ll_tab = list()
for (i in seq_along(cc$individuals)) {
  ll_tab[[i]] = c(i, cc$individuals[[i]]$n)
}
tab = do.call(rbind, ll_tab)
tab = as.data.frame(rbind(tab, colSums(tab)))
colnames(tab) = c("Server", cc$individuals[[1]]$bin)
tab0 = tab
for (j in seq_along(tab)[-1]) {
  tab[[j]] = paste0("$", ifelse(tab[[j]] < 5, tab[[j]], paste0("\\bm{", tab[[j]], "}")), "$")
}
tab[[1]] = paste0("$", tab[[1]], "$")
tab[6, 1] = "$\\sum$"

# LaTeX Table:
writeLines(knitr::kable(tab, format = "latex", escape = FALSE),
  con = here::here("tables/tab-cc.tex"))

knitr::kable(tab0)



# Visualize distributed calibration curve vs. pooled one:

gg_cal_pooled = plotCalibrationCurve(cc, size = 1.5, individuals = FALSE) +
    geom_line(data = cc_pooled, aes(x = prob, y = truth), color = "red")
gg_cal_pooled
```

```{r, include=FALSE}
# Save figure for paper
gg_cal_pooled_pub = plotCalibrationCurve(cc, size = 1, individuals = FALSE) +
    geom_line(data = cc_pooled, aes(x = prob, y = truth), color = "red") +
  scale_color_uchicago() +
  scale_fill_uchicago() +
  my_theme +
  theme(
    plot.title = element_text(size = 10),
    plot.subtitle = element_text(size = 9),
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 9),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    panel.border = element_rect(colour = "black", fill = NA, size = 0.5)
  ) +
  ggtitle("", "")

ggsave(plot = gg_cal_pooled_pub,
  filename = here::here("figures/analysis-distr-cal-pooled.pdf"),
  width = 0.5 * textwidth,
  height = textwidth * 0.5,
  units = "mm")

# evince(here::here("figures/analysis-distr-cal-pooled.pdf"))

gg_roc_pooled_pub = plot(roc_glm, plot_ci = FALSE) +
  geom_line(data = plt_emp_roc_data, aes(x = FPR, y = TPR), color = "red") +
  geom_errorbarh(data = df_auc, aes_string(y = "0.1", xmin = "lower", xmax = "upper"), height = 0.05) +
  geom_point(data = df_auc, aes_string(y = "0.1", x = "auc"), size = 3) +
  geom_point(data = df_auc, aes_string(y = "0.1", x = "auc"), size = 1, color = "white") +
  # annotate("text", x = df_auc$auc, y = 0.1, label = "AUC", vjust = -1, size = 3, family = font) +
  # annotate("text", x = df_auc$auc, y = 0.1, label = round(df_auc$auc, 2), vjust = 2, size = 2.5, family = font) +
  # annotate("text", x = df_auc$lower, y = 0.1, label = round(df_auc$lower, 2), hjust = 1.2, size = 3, family = font) +
  # annotate("text", x = df_auc$upper, y = 0.1, label = round(df_auc$upper, 2), hjust = -0.2, size = 3, family = font) +
  scale_color_uchicago() +
  scale_fill_uchicago() +
  my_theme +
  theme(
    plot.title = element_text(size = 10),
    plot.subtitle = element_text(size = 9),
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 9),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    panel.border = element_rect(colour = "black", fill = NA, size = 0.5)
  )

if (all(! grepl(font, x = ft$FamilyName))) {
  gg_roc_pooled_pub = gg_roc_pooled_pub +
    annotate("text", x = df_auc$auc, y = 0.1, label = "AUC", vjust = -1, size = 3) +
    annotate("text", x = df_auc$auc, y = 0.1, label = round(df_auc$auc, 2), vjust = 2, size = 2.5) +
    annotate("text", x = df_auc$lower, y = 0.1, label = round(df_auc$lower, 2), hjust = 1.2, size = 3) +
    annotate("text", x = df_auc$upper, y = 0.1, label = round(df_auc$upper, 2), hjust = -0.2, size = 3)
} else {
  gg_roc_pooled_pub = gg_roc_pooled_pub +
    annotate("text", x = df_auc$auc, y = 0.1, label = "AUC", vjust = -1, size = 3, family = font) +
    annotate("text", x = df_auc$auc, y = 0.1, label = round(df_auc$auc, 2), vjust = 2, size = 2.5, family = font) +
    annotate("text", x = df_auc$lower, y = 0.1, label = round(df_auc$lower, 2), hjust = 1.2, size = 3, family = font) +
    annotate("text", x = df_auc$upper, y = 0.1, label = round(df_auc$upper, 2), hjust = -0.2, size = 3, family = font)
}

ggsave(plot = gg_roc_pooled_pub,
  filename = here::here("figures/analysis-distr-roc-glm-pooled.pdf"),
  width = 0.5 * textwidth,
  height = textwidth * 0.5,
  units = "mm")

# evince(here::here("figures/analysis-distr-roc-glm-pooled.pdf"))

# Log:
logVector = function(x, sep = " - ") paste(x, collapse = sep)
log = data.frame(
  time = as.character(Sys.time()),
  auc_roc = roc_glm$auc,
  auc_emp = auc,
  ci_roc_lower = roc_glm$ci[1],
  ci_emp_lower = ci_emp[1],
  ci_roc_upper = roc_glm$ci[2],
  ci_emp_upper = ci_emp[2],
  brier_dist = brier,
  brier_pooled = brier_pooled,
  n1 = ddim[[1]][1],
  n2 = ddim[[2]][1],
  n3 = ddim[[3]][1],
  n4 = ddim[[4]][1],
  n5 = ddim[[5]][1],
  calibration_preds = logVector(cc$aggregated$prob),
  calibration_truths = logVector(cc$aggregated$truth),
  calibration_n = logVector(tail(tab0, 1)[-1]),
  calibration_nas = logVector(paste0(
    tail(tab0, 1)[-1] * cc$aggregated$missing_ratio, "/", tail(tab0, 1)[-1]))
)
if (! file.exists(here::here("data/log.csv"))) {
  write.csv(log, file = here::here("data/log.csv"), row.names = FALSE)
} else {
  log_old = read.csv(here::here("data/log.csv"))
  log = rbind(log, log_old)
  write.csv(log, file = here::here("data/log.csv"), row.names = FALSE)
}
```

```{r}
# Summary of the results used in the paper:
tex_results = rbind(
  data.frame(command = "\\cidistlower", value = round(roc_glm$ci[1], 4)),
  data.frame(command = "\\cidistupper", value = round(roc_glm$ci[2], 4)),
  data.frame(command = "\\ciemplower", value = round(ci_emp[1], 4)),
  data.frame(command = "\\ciempupper", value = round(ci_emp[2], 4)),
  data.frame(command = "\\aucdist", value = round(roc_glm$auc, 4)),
  data.frame(command = "\\aucpooled", value = round(auc, 4)),
  data.frame(command = "\\rocglmparamOne", value = round(roc_glm$parameter[1], 4)),
  data.frame(command = "\\rocglmparamTwo", value = round(roc_glm$parameter[2], 4)),
  data.frame(command = "\\bsemp", value = round(brier_pooled, 4)),
  data.frame(command = "\\ts", value = 2 * 365),
  data.frame(command = "\\nOne", value = ddim[[1]][1]),
  data.frame(command = "\\nTwo", value = ddim[[2]][1]),
  data.frame(command = "\\nThree", value = ddim[[3]][1]),
  data.frame(command = "\\nFour", value = ddim[[4]][1]),
  data.frame(command = "\\nFive", value = ddim[[5]][1]),
  data.frame(command = "\\privparOne", value = epsilon),
  data.frame(command = "\\privparTwo", value = delta),
  data.frame(command = "\\ltwosensUC", value = round(l2s, 4))
)
writeLines(paste0("\\newcommand{", tex_results[[1]], "}{", tex_results[[2]], "}"),
  here::here("tables/tab-results.tex"))
```

## Log out from DataSHIELD servers

```{r, output=FALSE}
datashield.logout(conn)
```
