---
title: DataSHIELD Use-case
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo=FALSE, message=FALSE, warning=FALSE}
source(here::here("R/generate-data.R"))
source(here::here("R/create-model.R"))
source(here::here("R/upload-data.R"))
source(here::here("R/install-ds-packages.R"))


knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figures/"
)

library(ggplot2)
library(ggsci)

### Theme setup:
## ==================================================================

# Set font (if available):
font = "Tinos"
sysfonts::font_add_google(font, font)
extrafont::loadfonts()
ft = extrafont::fonttable()

if (all(! grepl(font, x = ft$FamilyName))) stop("Cannot load font")

# to determine width, use the latex package `layouts` and put
# `\printinunitsof{mm}\prntlen{\textwidth}` into the tex file.
textwidth = 148.92

## Helper functions:
## =====================================================================

source(here::here("R/helper.R"))

#' Open pdf files using evince on linux machines
#' @param file to the pdf
evince = function(file) system(paste0("evince ", file, " &"))
```

<!-- README.md is generated from README.Rmd. Please edit that file -->

## About the repository

This repository contains a short use-case base on the three packages `ds.predict.base`, `ds.calibration`, and `ds.roc.glm`. The main intend is to have a use-case to demonstrate how to distributively evaluate a model using the distributed ROC-GLM.

The following contains the preparation of test data and a test model as "setup" while the second part is the analysis. .

### Structure of the repository

TODO

## Setup

### Install packages

Install all packages locally and also on the DataSHIELD test machine:

```{r}
remotes::install_github("difuture-lmu/ds.predict.base")
remotes::install_github("difuture-lmu/ds.calibration")
remotes::install_github("difuture-lmu/ds.roc.glm")
```

### Generate data and fit model

The used data is the `GBSG2` from the `TH.data` packages. For further details see the help page `?TH.data::GBSG2`. The task is to predict whether hormonal therapy shows an improvement w.r.t. survival time. The model we are using is a random forest from the `ranger` package. The following code uses the `GBSG2` data, splits it into train and test data with 60 % for training and 40 % for testing. The test data is further split into 5 parts that are uploaded to DataSHIELD and used to simulate the distributed setup.

```{r, eval=FALSE}
source(here::here("R/generate-data.R"))
source(here::here("R/create-model.R"))
```

### Install package on DataSHIELD and upload data

```{r, eval=FALSE}
source(here::here("R/upload-data.R"))
source(here::here("R/install-ds-packages.R"))
```

## Analysis

### Log into DataSHIELD test server

```{r}
library(DSI)
library(DSOpal)
library(dsBaseClient)

library(ds.predict.base)
library(ds.calibration)
library(ds.roc.glm)

library(ggplot2)

builder = newDSLoginBuilder()

surl     = "https://opal-demo.obiba.org/"
username = "administrator"
password = "password"

datasets = paste0("SRV", seq_len(5L))
for (i in seq_along(datasets)) {
  builder$append(
    server   = paste0("ds", i),
    url      = surl,
    user     = username,
    password = password,
    table    = paste0("DIFUTURE-TEST.", datasets[i])
  )
}

## Get data of the servers:
conn = datashield.login(logins = builder$build(), assign = TRUE)
datashield.symbols(conn)
```

### Push and predict

```{r}
## Load the pre-calculated logistic regression:
load(here::here("data/mod.Rda"))

## Push the model to the servers:
pushObject(conn, obj = mod)
datashield.symbols(conn)

## Predict the model on the data sets located at the servers:
predictModel(conn, mod, "probs", predict_fun = "ranger:::predict.ranger(mod, data = D)$predictions")
datashield.symbols(conn)
```

### Analyse calibration of the predictions

```{r}
brier = dsBrierScore(conn, "D$cens", "probs")
brier

cc = dsCalibrationCurve(conn, "D$cens", "probs")
gg_cal = plotCalibrationCurve(cc, size = 1)
gg_cal
```

```{r, include=FALSE}
# Save figure for paper
gg_cal_pub = gg_cal +
  scale_color_uchicago() +
  scale_fill_uchicago() +
  theme_minimal(base_family = font) +
  theme(
    plot.title = element_text(size = 10),
    plot.subtitle = element_text(size = 9),
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 9),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    panel.border = element_rect(colour = "black", fill = NA, size = 0.5)
  )

ggsave(plot = gg_cal_pub,
  filename = here::here("figures/analysis-distr-calibration.pdf"),
  width = 0.7 * textwidth,
  height = textwidth * 0.5,
  units = "mm")

# evince(here::here("figures/analysis-distr-calibration.pdf"))
```

### Evaluate the model using ROC analysis

```{r}
roc_glm = dsROCGLM(conn, "D$cens", "probs")
gg_distr_roc = plot(roc_glm)
gg_distr_roc
```

```{r, include=FALSE}
# Save figure for paper
df_auc = data.frame(lower = roc_glm$ci[1], upper = roc_glm$ci[2], auc = roc_glm$auc)

gg_roc_pub = plot(roc_glm, plot_ci = FALSE) +
  geom_errorbarh(data = df_auc,
    aes_string(y = "0.1", xmin = "lower", xmax = "upper"), height = 0.05) +
  geom_point(data = df_auc, aes_string(y = "0.1", x = "auc"), size = 3) +
  geom_point(data = df_auc, aes_string(y = "0.1", x = "auc"), size = 1, color = "white") +
  annotate("text", x = df_auc$auc, y = 0.1, label = "AUC", vjust = -1, size = 3) +
  annotate("text", x = df_auc$auc, y = 0.1, label = round(df_auc$auc, 2), vjust = 2, size = 2.5) +
  annotate("text", x = df_auc$lower, y = 0.1, label = round(df_auc$lower, 2), hjust = 1.2, size = 3) +
  annotate("text", x = df_auc$upper, y = 0.1, label = round(df_auc$upper, 2), hjust = -0.2, size = 3) +
  scale_color_uchicago() +
  scale_fill_uchicago() +
  theme_minimal(base_family = font) +
  theme(
    plot.title = element_text(size = 10),
    plot.subtitle = element_text(size = 9),
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 9),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    panel.border = element_rect(colour = "black", fill = NA, size = 0.5)
  )

ggsave(plot = gg_roc_pub,
  filename = here::here("figures/analysis-distr-roc-glm.pdf"),
  width = 0.5 * textwidth,
  height = textwidth * 0.5,
  units = "mm")

# evince(here::here("figures/analysis-distr-roc-glm.pdf"))
```


## Cross check on pooled test data

```{r}
#' Calculate TPR and FPRs to plot the empirical ROC curve
#'
#' @param labels (`integer()`) True labels as 0-1-coded vector.
#' @param scores (`numeric()`) Score values.
#' @return (`data.frame()`) of the TPR and FPRs.
simpleROC = function(labels, scores) {
  labels = labels[order(scores, decreasing = TRUE)]
  data.frame(
    TPR = cumsum(labels) / sum(labels),
    FPR = cumsum(! labels) / sum(! labels), labels)
}

# Load pooled test data and predict:
dat_test = read.csv(here::here("data/data-test.csv"), stringsAsFactors = TRUE)
probs = predict(mod, data = dat_test)$predictions

# Calculate empirical AUC and compare with distributed ROC-GLM
auc = pROC::auc(dat_test$cens, probs)
c(auc_emp = auc, auc_distr_roc_glm = roc_glm$auc)

# Calculate TPR and FPR values and add to distributed ROC-GLM plot
plt_emp_roc_data = simpleROC(dat_test$cens, probs)

gg_roc_pooled = plot(roc_glm) +
  geom_line(data = plt_emp_roc_data, aes(x = FPR, y = TPR, color = "Empirical"))
gg_roc_pooled

# Calculate pooled brier score and calibration curve
brier_pooled = mean((dat_test$cens - probs)^2)
c(brier_pooled = brier_pooled, brier_distr = brier)

cc_pooled = calibrationCurve("dat_test$cens", "probs", nbins = 10)

# Visualize distributed calibration curve vs. pooled one:

gg_cal_pooled = plotCalibrationCurve(cc, size = 1.5, individuals = FALSE) +
    geom_line(data = cc_pooled, aes(x = prob, y = truth), color = "red")
gg_cal_pooled
```

```{r, include=FALSE}
# Save figure for paper
gg_cal_pooled_pub = plotCalibrationCurve(cc, size = 1, individuals = FALSE) +
    geom_line(data = cc_pooled, aes(x = prob, y = truth), color = "red") +
  scale_color_uchicago() +
  scale_fill_uchicago() +
  theme_minimal(base_family = font) +
  theme(
    plot.title = element_text(size = 10),
    plot.subtitle = element_text(size = 9),
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 9),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    panel.border = element_rect(colour = "black", fill = NA, size = 0.5)
  )

ggsave(plot = gg_cal_pooled_pub,
  filename = here::here("figures/analysis-distr-cal-pooled.pdf"),
  width = 0.5 * textwidth,
  height = textwidth * 0.5,
  units = "mm")

# evince(here::here("figures/analysis-distr-cal-pooled.pdf"))

gg_roc_pooled_pub = plot(roc_glm, plot_ci = FALSE) +
  geom_line(data = plt_emp_roc_data, aes(x = FPR, y = TPR), color = "red") +
  geom_errorbarh(data = df_auc, aes_string(y = "0.1", xmin = "lower", xmax = "upper"), height = 0.05) +
  geom_point(data = df_auc, aes_string(y = "0.1", x = "auc"), size = 3) +
  geom_point(data = df_auc, aes_string(y = "0.1", x = "auc"), size = 1, color = "white") +
  annotate("text", x = df_auc$auc, y = 0.1, label = "AUC", vjust = -1, size = 3, family = font) +
  annotate("text", x = df_auc$auc, y = 0.1, label = round(df_auc$auc, 2), vjust = 2, size = 2.5, family = font) +
  annotate("text", x = df_auc$lower, y = 0.1, label = round(df_auc$lower, 2), hjust = 1.2, size = 3, family = font) +
  annotate("text", x = df_auc$upper, y = 0.1, label = round(df_auc$upper, 2), hjust = -0.2, size = 3, family = font) +
  scale_color_uchicago() +
  scale_fill_uchicago() +
  theme_minimal(base_family = font) +
  theme(
    plot.title = element_text(size = 10),
    plot.subtitle = element_text(size = 9),
    axis.text = element_text(size = 7),
    axis.title = element_text(size = 9),
    legend.title = element_text(size = 10),
    legend.text = element_text(size = 9),
    panel.border = element_rect(colour = "black", fill = NA, size = 0.5)
  )

ggsave(plot = gg_roc_pooled_pub,
  filename = here::here("figures/analysis-distr-roc-glm-pooled.pdf"),
  width = 0.5 * textwidth,
  height = textwidth * 0.5,
  units = "mm")

# evince(here::here("figures/analysis-distr-roc-glm-pooled.pdf"))
```


## Log out from DataSHIELD servers

```{r}
datashield.logout(conn)
```
